Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-18 paper/sections/acknowledgements.tex
[ paper ] lots of grammar fixes.

We thank Universidad EAFIT in Medell\'in, Colombia for funding
support. This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.

We thank Joe Leslie-Hurd for his support and comments about \Metis.
We also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module.

Last but not least, we thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting us to be part of the \Agda Implementors’ Meeting XXV.
We honored to present there part of this research.

>>>>

I would like to thank to Universidad EAFIT in Medell\'in, Colombia for funding
support of this research and my master studies.
This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.
My special thanks to Andrés Sicard-Ramiréz for its patience, valuable comments
during this research, and in general advices and discussions around many
topics of my interest.

I thank Joe Leslie-Hurd for his support and comments about \Metis.
I also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module.

Last but not least, we thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting us to be part of the \Agda Implementors’ Meeting XXV.
I was honored to present there part of this research.

### 2017-10-18 paper/sections/conclusions.tex
[ paper ] lots of grammar fixes.

We presented a reliable proof-reconstruction in type theory of \Metis
solutions for problems in classical propositional logic without
equality. As a product of this research, we built a proof-reconstruction tool written in \Haskell. This tool generates
\Agda proof-terms of \Metis \TSTP derivations. \Agda files
delivered by the aforementioned tool are supported by the
formalization in \Agda~\cite{AgdaProp,AgdaMetis} of the \Metis rules
described in Section~\ref{ssec:emulating-inferences}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic. This decision was first a
difficulty since it demands more detailed proof-objects, a
description of every transformation or deduction step performed by
the prover, see
Section~\ref{sssec:normalization} and Section~\ref{sssec:reordering}.
But, we chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

One of the main contribution of this study was
increasing the trustworthiness of the automatic provers,
in our case, \Metis.
Justifying a proof-object of a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these system (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}} in \Metis.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

>>>>

We presented a reliable proof-reconstruction in type theory of \Metis
solutions for problems in classical propositional logic without
equality. As a product of this research, we were able to build a proof-reconstruction tool written in \Haskell. This tool generates
\Agda proof-terms of \Metis \TSTP derivations. \Agda files
delivered by the aforementioned tool are supported by the
formalization in \Agda~\cite{AgdaProp,AgdaMetis} of the \Metis rules
described in Section~\ref{ssec:emulating-inferences}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proof-objects, a
description of every transformation or deduction step performed by
the prover, see
Section~\ref{sssec:normalization} and Section~\ref{sssec:reordering}.
But, we chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

One of the main contribution of this study was
increasing the trustworthiness of the automatic provers,
in our case, \Metis.
Justifying a proof-object of a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}} in \Metis.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

### 2017-10-18 paper/sections/introduction.tex
[ paper ] lots of grammar fixes.

Proof-reconstruction is a hard labor since it depends on the
integration of two complex systems. On the one hand, we have the
automatic theorem provers (henceforth \ATP) and their specification
logic. These tools are usually classified in at least one of the
following categories.
A \SAT solver (e.g., \name{zChaff}~\cite{Moskewicz2001} and
\name{MiniSat}~\cite{Een2004}) to prove unsatisfiability of \CNF
formulas, a \abbre{QBF} solver (e.g.,
\name{GhostQ}~\cite{Klieber2014} and
\name{DepQBF}~\cite{Lonsing2017}) to prove satisfiability and
invalidity of quantified Boolean formulas, a \SMT solver
(e.g \name{CVC4}~\cite{Barrett2011},\name{veriT}~\cite{bouton2009},
and \name{Z3}~\cite{DeMoura2008}) to prove unsatisfiability of
formulas from first-order logic with theories, and a prover for
validity of formulas from first-order logic with equality
(e.g.,\name{E}~\cite{Schulz:AICOM-2002},
\name{leanCoP}~\cite{Otten2008},
\Metis~\cite{hurd2003first},
\name{SPASS}~\cite{Weidenbach2009} and
\name{Vampire}~\cite{Riazanov1999}), high-order logic (e.g.,
\name{Leo-II}~\cite{Benzmuller2008} and
\name{Satallax}~\cite{Brown2012}) or intuitionistic logic (e.g.,
\name{ileanCoP}~\cite{Otten2008},
\name{JProver}~\cite{Schmitt2001}, and
\name{Gandalf}~\cite{Tammet1997}), among others. On the other hand,
we have the proof checkers, interactive theorem provers (henceforth
\ITP) or proof assistants (e.g., \Agda~\cite{agdateam},
\name{Coq}~\cite{coqteam},
\name{Isabelle}~\cite{paulson1994isabelle}, and
\name{HOL4}~\cite{norrish2007hol}). The \ITP tools provide us the
logic framework to check and validate the reply of the \ATPs since
they allow us to define the formal language for the problems like
operators, logic variables, axioms, and theorems.

A proof-reconstruction tool provides such an integration translating
the reply delivered by the prover into the formalism of the proof
assistant. Because the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the
target, the reconstruction turns out in a ``reverse engineering''
task. Then, reconstructing a proof involves an in-depth
understanding of the algorithms in the \ATP and the specification
logic in the \ITP. To begin with the proof-reconstruction, it is
necessary to have a proof-object from the \ATP in a consistent
format to work with, that is, a full script describing step-by-step
with exhaustive details and without ambiguities, the derivation to
get the actual proof.
For problems in classical propositional logic (henceforth \CPL),
from a list of at least forty\footnote{\ATPs available from the web
service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
provers were able to deliver proofs (e.g.,
\name{CVC4}~\cite{Barrett2011}, \name{SPASS}, and
\name{Waldmeister}~\cite{hillenbrand1997}) and a little bit less replied with
a proof
in a file format like \TSTP~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}
(e.g., \name{E}, \Metis, \name{Vampire}, and \name{Z3}),
\len{LFSC}~\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

@@

In this paper, we ported a subset of the \Metis' inference rules to
\Agda, the propositional fragmented without equality. The formalization of
these rules allowed us to justify step-by-step \Metis' \TSTP derivations by
translating automatically to \Agda by a proof-reconstruction tool built for
this research named \Athena.
This paper has been organized in the following way.
The first section of the paper, Section~\ref{ssec:CPL}, will mention some
notation and syntax of the classical propositional logic.
In Section~\ref{ssec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we present our
approach to reconstruct proofs deliver by \Metis in type theory.
Section~\ref{sec:conclusions} summarizes the main findings and includes a
discussion of future work.
Some related works with this paper is presented in Section~\ref{sec:related-work}.
Finally, we provide a small example in the appendix \textsc{A} of the
proof-reconstruction workflow.

>>>>

Proof-reconstruction is a hard labor since it depends on the
integration of two complex systems. On the one hand, we have the
automatic theorem provers (henceforth \ATP) and their specification
logic. These tools are usually classified in at least one of the
following categories.
A \SAT solver (e.g., \name{zChaff}~\cite{Moskewicz2001} and
\name{MiniSat}~\cite{Een2004}) to prove unsatisfiability of \CNF
formulas, a \abbre{QBF} solver (e.g.,
\name{GhostQ}~\cite{Klieber2014} and
\name{DepQBF}~\cite{Lonsing2017}) to prove satisfiability and
invalidity of quantified Boolean formulas, a \SMT solver
(e.g \name{CVC4}~\cite{Barrett2011}, \name{veriT}~\cite{bouton2009},
and \name{Z3}~\cite{DeMoura2008}) to prove unsatisfiability of
formulas from first-order logic with theories, and a prover for
validity of formulas from first-order logic with equality
(e.g., \name{E}~\cite{Schulz:AICOM-2002},
\name{leanCoP}~\cite{Otten2008},
\Metis~\cite{hurd2003first},
\name{SPASS}~\cite{Weidenbach2009} and
\name{Vampire}~\cite{Riazanov1999}), high-order logic (e.g.,
\name{Leo-II}~\cite{Benzmuller2008} and
\name{Satallax}~\cite{Brown2012}) or intuitionistic logic (e.g.,
\name{ileanCoP}~\cite{Otten2008},
\name{JProver}~\cite{Schmitt2001}, and
\name{Gandalf}~\cite{Tammet1997}), among others. On the other hand,
we have the proof checkers, interactive theorem provers (henceforth
\ITP) or proof assistants (e.g., \Agda~\cite{agdateam},
\name{Coq}~\cite{coqteam},
\name{Isabelle}~\cite{paulson1994isabelle}, and
\name{HOL4}~\cite{norrish2007hol}). The \ITP tools provide us the
logic framework to check and validate the reply of the \ATPs since
they allow us to define the formal language for the problems i.e.,
operators, logic variables, axioms, and theorems.

A proof-reconstruction tool provides such an integration, translating
the reply generated by the prover into the formalism of the proof
assistant. Since the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the
target, the reconstruction turns out to be a ``reverse engineering''
task. Then, reconstructing a proof involves an in-depth
understanding of the algorithms in the \ATP and the logic specification
in the \ITP. To begin with the proof-reconstruction, it is
necessary to have a proof-object from the \ATP in a consistent
format, that is, a full script describing step-by-step
with exhaustive details and without ambiguities, the derivation to
get the actual proof.
For problems in classical propositional logic (henceforth \CPL),
from a list of at least forty\footnote{\ATPs available from the web
service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
provers were able to deliver proofs (e.g.,
\name{CVC4}~\cite{Barrett2011}, \name{SPASS}, and
\name{Waldmeister}~\cite{hillenbrand1997}) and fewer replied with
a proof
in a file format like \TSTP~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}
(e.g., \name{E}, \Metis, \name{Vampire}, and \name{Z3}),
\len{LFSC}~\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

@@

In this paper, we ported a subset of the \Metis' inference rules to
\Agda, the propositional fragmented without equality. The formalization of
these rules allowed us to justify step-by-step \Metis' \TSTP derivations by
translating automatically to \Agda by a proof-reconstruction tool built for
this research named \Athena.
This paper has been organized in the following way.
The first section of the paper, Section~\ref{ssec:CPL}, will mention some
notation and syntax of the classical propositional logic.
In Section~\ref{ssec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we present our
approach to reconstruct proofs deliver by \Metis in type theory.
Section~\ref{sec:conclusions} summarizes the main findings and includes a
discussion of future work.
Some related works to this paper are presented in Section~\ref{sec:related-work}.
Finally, in appendix \textsc{A}, we provide a small example of the
proof-reconstruction workflow.

### 2017-10-18 paper/sections/proof-reconstruction.tex
[ paper ] lots of grammar fixes.

\begin{theorem}[thm-conjunct]
  \label{thm:thm-conjunct}
  $Γ ⊢ φ$, and $ψ  : \Prop$ , then $Γ ⊢ conjunct(φ, ψ)$.
\end{theorem}
\begin{proof}\hspace{2cm}
\begin{itemize}
  \item For the case $φ ≡ ψ$, $Γ ⊢ conjunct(φ, ψ)$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $subst$ theorem.
  \item If the proposition $φ$ is a conjunction, and we can get $ψ ≡ conjunct(φ_{i}, ψ)$ for some $i = 1,\ 2$, then,

@@

\begin{itemize}
  \item Reorder a disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$ to match with
$ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$. For all $i$, there exists at least one $j$ such that $φᵢ ≡ ψⱼ$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$. For all $i$, there exists at least one $j$ such that $φᵢ ≡ ψⱼ$.
\end{itemize}

First, we define the $rdisj$ function to convert
a disjunction in its right-associative form.
The $rconj$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

@@

The $build_{∨}$ function in Eq.~(\ref{eq:build-or}) intends to construct a
disjunction from another disjunction, specifically, this functions will try
to rearrange the disjuncts in the source formula to match with the target
disjunction formula, similar as we describe above.

@@

From now, we assume all propositions to be right-associative unless otherwise stated.
The $factor$ function in Eq.~(\ref{eq:factor-definition}) intends to simplify a
special case of disjunction, the repeated disjuncts (e.g., $factor(φ ∨ φ) \mapsto\ φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $(ψ ∨ φ)$.

@@

To construct a disjunction $ψ$ from a disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove equality
between nested disjunctions. The result is the $sbuild_{∨}$ function that uses
every disjunct from the source formula, $φ$, to build up the target disjunction $ψ$.

@@

\begin{example}
Let see the main idea behind the lem-sbuild$_{∨}$ lemma.
We build the disjunction $((p ∨ q) ∨ r)$ from the disjunction
$(r ∨ (q ∨ p))$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ q$}
  \RightLabel{∨-intro₂}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \RightLabel{∨-intro₁}
  \UnaryInfC{$Γ ⊢ (p ∨ q) ∨ r$}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder_{∧} : \Prop → \Prop \to \Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-reorder$_{∧}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ reorder_{∧}(φ, ψ)$.
\end{lemma}

@@

To emulate this rule, we perform a sequence of rearrangements inside the
formulas involved to match with the pattern expected by the resolution theorem
described above. We begin by defining a convenient version of
the binary resolution in Def.~\ref{def:rsol}.

@@

In Def.~\ref{def:resolve}, the emulated rule for $resolve$
has four arguments. The first three were described
above, the last one, ψ, represents the expected conclusion in the \Metis
derivation.

@@

In this section we provide functions and theorems to remove, inside a normalized formula, the definitions showed in
Fig.~\ref{fig:conjunctive-disjunctive-simpl}.

@@

Now, we have removed redundancies in the disjunctions by applying
$rm_{∨}$ function, we can define a similar function to work with
conjunctions, i.e., $rm_{∧}$.

@@

Continuing with the rest of tautologies from
Fig.~\ref{fig:conjunctive-disjunctive-simpl}, we define the
$ndisj_{∨}$ function and $ndisj_{∨}$ function that
take into account the following cases.

@@

Finally, we can translate formulas from \Prop to \NProp by defining
the $nnf₀$ function. This function is based on the $nnf$ function described in
\citeauthor{Bezem2002}~\cite{Bezem2002}. Their $nnf$ function
translates formulas to their classical negative normal form. In the
reference, the authors avoid a termination problem with a similar
function by using, as an extra argument, the polarity of the formula.
We follow the same idea in Def.~\ref{def:nnf-zero}.
Therefore, we need the \abbre{Polarity} type with
two constructors: $⊕$ to denote positive polarity, and $⊖$ to denote
negative polarity of the formula.

@@

The \canonicalize rule is an overloaded inference rule which performs
normalization of the proposition input.
%This rule can introduce axioms and definitions into the problem solution.
This rule converts a \texttt{fof} %TODO
formula in clausal form to a \CNF clause but also it performs some
simplifications in the process (e.g., removing ⊤ and ⊥ in the formula).

@@

In the following \Metis \TSTP derivation, we see an example where
\clausify expands the formula \texttt{normalize\_0\_0} by using distributive
laws presented in conjunctive normal form conversion.

@@

\begin{remark}
Besides the fact that $\List\ \Prop \to \Prop$ is the type that coincides
with the \simplify rule in \TSTP derivations, we choose a different
option. In the translation from \TSTP to \Agda, we take the list of
derivations and use a classical left folding ($foldl$) with
$simplify$ function over the list of $φ₁, φ₂, \cdots, φₙ$
that avoids us to define a new sequent type to
support \List \Prop type in the conclusion side.
\end{remark}

>>>>

\begin{theorem}[thm-conjunct]
  \label{thm:thm-conjunct}
  $Γ ⊢ φ$, and $ψ  : \Prop$ , then $Γ ⊢ conjunct(φ, ψ)$.
\end{theorem}
\begin{proof}\hspace{2cm}
\begin{itemize}
  \item For the case, $φ ≡ ψ$, $Γ ⊢ conjunct(φ, ψ)$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $subst$ theorem.
  \item If the proposition $φ$ is a conjunction, and we can get $ψ ≡ conjunct(φ_{i}, ψ)$ for some $i = 1,\ 2$, then,

@@

\begin{itemize}
  \item Reorder a disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$ to match with
$ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$, when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$ when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
\end{itemize}

First, we define the $rdisj$ function to convert
a disjunction in its right-associative form.
The $rconj$ function converts conjunction to its right-associative form is not
presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

@@

The $build_{∨}$ function in Eq.~(\ref{eq:build-or}) intends to construct a
disjunction from another disjunction, specifically, this functions will try
to rearrange the disjuncts in the source formula to match with the target
disjunction formula, similar as we described above.

@@

From now, we assume all propositions are right-associative unless otherwise stated.
The $factor$ function in Eq.~(\ref{eq:factor-definition}) intends to simplify a
special case of disjunction, the repeated disjuncts (e.g., $factor(φ ∨ φ) \mapsto\ φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $(ψ ∨ φ)$.

@@

To construct a disjunction $ψ$ from another disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove equality
between nested disjunctions. The result is the $sbuild_{∨}$ function that uses
every disjunct from the source formula, $φ$, to build up the target disjunction $ψ$.

@@

\begin{example}
Let see the main idea behind the lem-sbuild$_{∨}$ lemma.
We will build the disjunction $((p ∨ q) ∨ r)$ from the disjunction
$(r ∨ (q ∨ p))$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ q$}
  \RightLabel{∨-intro₂}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \RightLabel{∨-intro₁}
  \UnaryInfC{$Γ ⊢ (p ∨ q) ∨ r$}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder : \Prop → \Prop \to \Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-reorder$_{∧}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ reorder(φ, ψ)$.
\end{lemma}

@@

To emulate this rule, we perform rearrangements inside the formulas
involved to match with the pattern expected by the resolution theorem
described above. We begin by defining a convenient version of
the binary resolution in Def.~\ref{def:rsol}.

@@

In Def.~\ref{def:resolve}, the emulated rule for $resolve$
has four arguments. The first three were described
above. The last one, ψ, represents the expected conclusion in the \Metis
derivation.

@@

In this section we provide the functions and theorems to
remove inside a normalized formula the definitions showed in
Fig.~\ref{fig:conjunctive-disjunctive-simpl}.

@@

Now, we have removed redundancies in the disjunctions by applying
$rm_{∨}$ function, we can define a similar function to work with
conjunctions, the $rm_{∧}$ function.

@@

Continuing with the rest of tautologies from
Fig.~\ref{fig:conjunctive-disjunctive-simpl}, we define
$ndisj_{∨}$ function and $ndisj_{∨}$ function that
take into account the following cases.

@@

Finally, we can translate formulas from \Prop to \NProp by defining
the $nnf₀$ function. This function is based on the $nnf$ function described in
\citeauthor{Bezem2002}~\cite{Bezem2002}. Their $nnf$ function
translates formulas to their classical negative normal form. In the
reference, the authors avoid a termination problem with a similar
function by using as an extra argument, the polarity of the formula.
We follow the same idea in Def.~\ref{def:nnf-zero}.
Therefore, we need the \abbre{Polarity} type with
two constructors: $⊕$ to denote positive polarity, and $⊖$ to denote
negative polarity of the formula.

@@

The \canonicalize rule is an overloaded inference rule which performs
normalization of the proposition input.
%This rule can introduce axioms and definitions into the problem solution.
This rule converts a \texttt{fof} %TODO
formula in clausal form to a \CNF clause but also it perform some
simplification in the process (e.g., removing ⊤ and ⊥ in the formula).

@@

In the following \Metis \TSTP derivation, we see a clear example where
\clausify expands the formula \texttt{normalize\_0\_0} by using distributive
laws presented in conjunctive normal form conversion.

@@

\begin{remark}
Besides the fact, $\List\ \Prop \to \Prop$ is the type that coincides
with the \simplify rule in \TSTP derivations, we choose a different
option. In the translation from \TSTP to \Agda, we take the list of
derivations and use a classical left folding ($foldl$) with
$simplify$ function over the list of $φ₁, φ₂, \cdots, φₙ$.
That avoids us to define a new sequent type to
support \List \Prop type in the conclusion side.
\end{remark}
